---
title: "Lecture 14: Other Models"
author: "Taylor Arnold"
date: " "
fontsize: 10pt
output:
  beamer_presentation:
    template: template.beamer
    pandoc_args: "--latex-engine=xelatex"
    theme: "metropolis"
    fonttheme: "professionalfonts"
    slide_level: 2
    df_print: tibble
    fig_width: 9
    fig_height: 5.5
    fig_caption: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)
library(dplyr)
options(dplyr.print_max = 6, dplyr.print_min = 6)
options(width = 60)

library(dplyr)
library(readr)
library(ggplot2)
library(smodels)
library(forcats)

theme_set(theme_minimal())
msleep <- na.omit(msleep)

set.seed(1)
```

## lm()

So far, we have seen how to use the function `lm_basic` to
build linear regression models.

As I probably mentioned before, this is just a soft wrapper
around the base R function `lm`, with the following tweaks:

- you are forced to supply a data frame
- the summary table produces confidence intervals rather
than standard errors

I don't use this function in my own work, but find it easier
for teaching.

## lm()

One major benefit of the "real" `lm` is that there are many
more options available for calling it, such as including
sample weights.

## glm()

There are many other functions in R that have similar calling
mechanisms to `lm` but run different underlying models.

For example, `glm` fits generalized linear models. These can
be used, amongst other things, for fitting a model to a binary
response:

```{r}
df <- data_frame(y = c(0,0,0,0,1,1,1,1), x = rnorm(8))
model <- glm(y ~ x, data = df, family = binomial())
```

## glm()

```{r}
summary(model)
```

## rlm()

In the **MASS** package (included with all standard R
installations) is the `rlm` function for fitting robust
linear regression:

```{r}
library(MASS)
x <- rnorm(100)
y <- 1 + x + rnorm(100, sd = 0.2)
y[50] <- 1e4
model_lm <- lm(y ~ x)
model_rlm <- rlm(y ~ x)
```

## rlm()

```{r}
summary(model_lm)
```

## rlm()

```{r}
summary(model_rlm)
```

## other models

If you have a need for a specific model, you can usually
find an R package that support it. In most cases, the model
will roughly resemble calling `lm`.

Some common examples you may run into:

- `gam::gam` for generalized additive models
- `nls` for non-linear regression
- `lme4::lmer` for mixed effects models
- `quantreg::qr` for quantila regression
- `glmnet::glmnet` for the generalized elastic net
- `randomforest::randomforest` for random forest classifications
- `forcast::auto.arima` for modeling time series





