---
title: "Lecture 06: Data Manipulation"
author: "Taylor Arnold"
date: " "
fontsize: 11pt
output:
  beamer_presentation:
    template: template.beamer
    pandoc_args: "--latex-engine=xelatex"
    theme: "metropolis"
    fonttheme: "professionalfonts"
    slide_level: 2
    df_print: tibble
    fig_width: 9
    fig_height: 5.5
    fig_caption: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)
library(dplyr)
options(dplyr.print_max = 6, dplyr.print_min = 6)
options(width = 60)

library(ggplot2)
library(dplyr)
data(msleep)
library(smodels)
```

## Data manipulation

- so far we have primarily been working with a single
dataset as it was directly given to us. In many cases it is
advantageous to produce new datasets from our original one.

- can be as simple as selecting a subset of the original columns
or rows, or as involved as changing the *level of analysis* of
the entire dataset.

- we will introduce several functions called *verbs* that assist
in manipulating datasets. These all come from the package **dplyr**,
supplemented with functions in **smodels**.

## select

In this handout, as with the others so far, I will use the `msleep`
dataset in order to show various numerical summaries. Rather than
the entire dataset, it will be helpful to take just a subset of
the columns. This gives us the first verb we'll see: `select`:

```{r}
msleep <- select(msleep, awake, vore, genus)
```

# A grammar for data manipulation

The *verbs* in **dplyr** take either one or
two existing datasets and return a new dataset. There are roughly 30
different verbs, of which we'll use only about 4 in this course:

* `filter`
* `mutate`
* `count`
* `group_summarize`
* `left_join`

These verbs on their own are relatively straightforward, but can be tied
together to produce surprisingly complex new datasets.

## Filtering rows

The filter function takes a dataset followed by a logical statements.
It returns a dataset that has any rows in the input data that match the
filtering statements. For example, the following returns a dataset for
all mammals that are awake more than 20 hours per day:

```{r echo = TRUE, eval = FALSE}
filter(msleep, awake > 20)
```

## Filtering rows

```{r echo = TRUE, eval = TRUE}
filter(msleep, awake > 20)
```

## Filtering rows

It is very important to notice that the original `msleep` dataset has not been
altered here. If we want to actually work with the filtered data,
we need to save it using the assignment operator `<-` and given it a name. To
pull out just the *sleepy* mammals, we could do the following:

```{r echo = TRUE, eval = FALSE}
sleepy <- filter(msleep, awake < 6)
```

## Filtering rows

You should see now that a new dataset appears in your workspace named `sleepy`.
It is possible to work with this new dataset in all of the ways we have plotted and
(now) filtered the original data.

## Constructing new variables

The `mutate` function preserves all rows of the original dataset, unlike `filter`,
but adds a new variables. For example, to add hours asleep into the dataset
we can do this:

```{r echo = TRUE}
msleep <- mutate(msleep, asleep = 24 - awake)
```

Here, I used the assignment command to save the result back into the `msleep`
dataset.

## Constructing new variables

If you look in your environment window, you'll still see just a single
version of `msleep` but this one will have one extra variables. This is a relatively
safe pratice with the mutate function, as we are simply adding information, but
should be generally avoided when using filter.

It is possible to use mutate to redefine an existing variable by giving mutate a
variable name that already exists. Be careful of this, particularly if you are
overwritting the original dataset.

## Constructing new variables

Of course, we could also do the same thing with the following:

```{r}
msleep$asleep <- 24 - msleep$awake
```

It is just nice to use the `mutate` function as it is consistent
with all of the other **dplyr** verbs.

## Summarizing data

The `group_summarize` function is, in my opinion, the most complex verb that
we will use this semester. If we use the function on a dataset without any
other options it gives the mean, median, standard deviation, and sum for
every numeric  variable in the dataset. An overall count is also included.
Let's apply it to `msleep`:

```{r echo = TRUE, eval = FALSE}
group_summarize(msleep)
```

## Summarizing data

```{r echo = TRUE, eval = TRUE}
group_summarize(msleep)
```

## Summarizing data

These variables could have easily been computed by calling the respective
functions individually in R. The group summarize function becomes more
interesting when we pass it a second input giving a variable to group by.
For example, here is the summary *grouped by vore*

```{r echo = TRUE, eval = FALSE}
group_summarize(msleep, vore)
```

## Summarizing data

```{r echo = TRUE, eval = TRUE}
group_summarize(msleep, vore)
```

## Summarizing data

Notice that the result now provides these summaries for each group. It
is possible to summarize by multiple groups at once, which produces
summaries for each unique combination of the those variables. For instance,
we could summarize by both genus and vore:

```{r echo = TRUE, eval = FALSE}
group_summarize(msleep, vore, genus)
```

Which returns a row for each unique combination of `vore` and `genus`.

## Summarizing data

```{r echo = TRUE, eval = TRUE}
group_summarize(msleep, vore, genus)
```

## Counting data

The summarized data also includes a column called `n` giving the total
number of observations within a group. The `count` function functions
similarly, but does not returns any of the other summary functions. This
is convenient if all you need are counts:

```{r echo = TRUE, eval = FALSE}
count(msleep, vore)
```

## Counting data

```{r echo = TRUE, eval = TRUE}
count(msleep, vore)
```

## Counting data

The `sort` option gives us the ability to order to data at the same time
as it is counted:

```{r echo = TRUE, eval = FALSE}
count(msleep, vore, sort = TRUE)
```

## Counting data

```{r echo = TRUE, eval = TRUE}
count(msleep, vore, sort = TRUE)
```

# Combining datasets

## left_join

The final verb that we will use this semester is also the only two-table
verb that we will need. It will be used to combine a dataset with metadata
about one or more of its variables. To illustrate, let's make a small
data frame that contains the full name for the short-hand abbreviations
given in the variable `vore`.

```{r echo = TRUE}
meta <- data_frame(
          vore = c("carni", "omni", "herbi", "insecti"),
         full_name = c("carnivore", "omnivore",
                       "herbivore", "insectivore"))
```

## left_join

```{r}
meta
```

## left_join

To combine these with the original dataset, we use the `left_join` function,
giving the larger dataset first:

```{r echo = TRUE}
msleep <- left_join(msleep, meta)
```

There is a new variable `full_name` that now displays the
full name for each `vore` type.

## left_join

```{r}
msleep
```



