---
title: "Lecture 25: Restructuring data"
author: "Taylor Arnold"
date: " "
fontsize: 10pt
output:
  beamer_presentation:
    template: template.beamer
    pandoc_args: "--latex-engine=xelatex"
    theme: "metropolis"
    fonttheme: "professionalfonts"
    slide_level: 2
    df_print: tibble
    fig_width: 9
    fig_height: 5.5
    fig_caption: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)
library(dplyr)
options(dplyr.print_max = 6, dplyr.print_min = 6)
options(width = 60)

library(dplyr)
library(readr)
library(ggplot2)
library(smodels)
library(forcats)

theme_set(theme_minimal())

set.seed(1)
```

## gathering data

In some cases we have datasets where multiple columns could be
treated as individual observations. What does that mean? Think
of the cancer dataset we used earlier in the semester, taking
off just the cancer incidence rates:

```{r}
cancer <- read_csv("https://statsmaths.github.io/stat_data/cancer_inc_data.csv")
cancer <- select(cancer, breast, colorectal, prostate, lung,
                 melanoma)
```

## gathering data

```{r}
cancer
```

## gathering data

A common question with this data is: how we could plot all of the
cancer types on the same plot. The canonical way of doing this
would be to make a new dataset where each row, instead of being
a single county, is then just one incidence rate. In other words
each county will have five rows associated with it. To do this
we use the `gather` function. It is found in the package **tidyr**.

## gathering data

```{r}
library(tidyr)
gather(cancer)
```

## gathering data

If we want to give the name of the `key' and `value', those are
given as the next two parameters to gather:

```{r}
gather(cancer, type, incidence)
```

## gathering data

In many cases, there are other variables that we want to be duplicated
along with the other keys. For example, take the speed skating dataset,
selecting off a few variables to make it a bit more tractable:

```{r}
speed <- read_csv("https://statsmaths.github.io/stat_data/speed_skate.csv")
speed <- select(speed, num_skater, nationality, time_lap1,
                time_lap2, time_lap3,
                time_lap4, time_lap5)
```

## gathering data

We can indicate the variables that should not be gathered by including
them after the key and value terms with minus signs:

```{r}
gather(speed, skater, value, -num_skater, -nationality)
```

## gathering data

Or, we can not use the minus sign and include on those variables
that should be gathered:

```{r}
gather(speed, skater, value, time_lap1, time_lap2,
        time_lap3, time_lap4, time_lap5)
```

## gathering data

The results are the same; in most cases one or the other will lead to
less typing. There is also a complement to gathering called
spreading, available through the function `spread`. You should not
need that in this class because every dataset is maximally spread
already. It is also a lot more difficult to use because you have to
be careful about implicit missing values and what to do with them.


