---
title: "Class 04: "
author: "Taylor Arnold"
output: html_notebook
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(eval = TRUE)
knitr::opts_chunk$set(fig.path = "../assets/2017-09-07-class04/")
knitr::opts_chunk$set(fig.height = 5)
knitr::opts_chunk$set(fig.width = 8.5)
knitr::opts_chunk$set(out.width = "100%")
knitr::opts_chunk$set(dpi = 300)
```

```{r, message = FALSE}
library(readr)
library(ggplot2)
library(dplyr)
library(viridis)
```

Another layer type with a default value are **scales**, which
are associated with every variable-assigned aesthetic. For
example, the `scale_x_continuous` and `scale_y_continuous`
functions define the range of the two dimensions of the plot.
The defaults (just fitting all of the data) usuall work well,
but we can modify them when needed:

```{r}
ggplot(mpg, aes(displ, hwy)) +
  geom_point() +
  scale_x_continuous(limits = c(0, 7.5))
```

There are also alternative scales for the x and y coordinates,
such as `scale_x_log10` and `scale_x_sqrt`.

The color scales determine the exact colors being used. The
**viridis** package provides an alternative scale for the
colors that I prefer:

```{r}
ggplot(mpg, aes(displ, hwy)) +
  geom_point(aes(color = cty)) +
  scale_color_viridis()
```

The viridis package is particularly useful when producing
plots for overhead projectors, plots that may need to printed,
or plots that are color-blind friendly.




